Estimating Neighborhood Rents with Scraped Web Data

### 2018-06-02 index.rmd
working on manuscript

---
title: "smooth-sea"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "hesscl.com", href: "http://hesscl.com", align: right }
    orientation: columns
    theme: bootstrap
    logo: logo.svg
    source_code: "https://github.com/hesscl/smooth-sea"
---

@@

#load ACS data extract (codebook in folder)
census <- read_csv(file = "./input/acsExtract.csv")

#load demo extract
#demo <- read_csv(file = "./input/tractDemoSumStats.csv")

#load culture extract
#culture <- read_csv(file = "./input/tractCultureSumStats.csv")

#load police report extract
#police <- read_csv(file = "./input/tractPoliceReportSumStats.csv")

@@

#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  #connect to database if file.exists
  DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
  cl <- tbl(DB, "clean") #create dbi for clean listing table

  #compute tract aggregates for CL listing count
  tract <- cl %>%
    collect %>% #bring db query into memory
    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
    mutate(listingDate = as.Date(listingDate),
           listingQtr = as.yearqtr(listingDate)) %>%
    filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
    group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
    summarize(nListings = n(),
              n1B = sum(cleanBeds == 1),
              medRent = median(cleanRent),
              med0B = median(cleanRent[cleanBeds==0]),
              med1B = median(cleanRent[cleanBeds==1]),
              med2B = median(cleanRent[cleanBeds==2]),
              med3B = median(cleanRent[cleanBeds==3]),
              lagRent = lag(cleanRent)) %>% #create tract aggregates
    ungroup %>%
    arrange(GISJOIN, listingQtr)
  dbDisconnect(DB)
  write_csv(tract, "./input/tractCl.csv")

} else{
  #if github config, read in extract from cl db
  tract <- read_csv("./input/tractCl.csv")
}

@@

#panel <- left_join(panel, demo)
#panel <- left_join(panel, culture)
#panel <- left_join(panel, police)
#panel <- left_join(panel, sea_shp@data)

sea_df <- panel %>% #set to NA for INLA to forecast
  mutate(actualRent = med1B,
         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>%
  mutate(p1B = n1B/nListings) %>%
  arrange(listingQtr, GISJOIN)

@@

This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.

>>>>

---
title: "smooth-sea"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "hesscl.com", href: "http://hesscl.com", align: right ,
          title: "Manuscript", href: "http://github.com/hesscl/smooth-sea/smooth-sea.html", align: right}
    orientation: columns
    theme: bootstrap
    logo: logo.svg
    source_code: "https://github.com/hesscl/smooth-sea"
---

@@

#source data_load()
source("helpers/load_data.R")

#load ACS data extract (codebook in folder)
census <- read_csv(file = "./input/acsExtract.csv")
censusRent <- read_csv(file = "./input/acsRent.csv")
census <- inner_join(census, censusRent)

@@

#use sourced function to load listing summaries
tract <- load_data()

@@

#establish object for modeling based off of adjusting `panel`
sea_df <- panel %>% 
  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
  arrange(listingQtr, GISJOIN) #arange qtr, tract

@@

This website provides interactive results for the forecasting models explored in the paper [_Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_](http://github.com/hesscl/smooth-sea/smooth-sea.html).

### 2018-06-02 smooth-sea.Rmd
working on manuscript

---
title: "smooth-sea"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "hesscl.com", href: "http://hesscl.com", align: right ,
          title: "Manuscript", href: "http://github.com/hesscl/smooth-sea/smooth-sea.html", align: right}
    orientation: columns
    theme: bootstrap
    logo: logo.svg
    source_code: "https://github.com/hesscl/smooth-sea"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
library(tidyverse) #need dev version of ggplot
library(forcats)
library(haven)
library(lubridate)
library(haven)
library(zoo)
library(ggthemes)
library(RColorBrewer)
library(sp)
library(spdplyr)
library(geosphere)
library(rgdal)
library(rgeos)
library(sqldf)
library(spdep)
library(viridis)
library(latex2exp)
library(plotly)
library(DT)
library(crosstalk)

#source data_load()
source("helpers/load_data.R")

@@

#use sourced function to load listing summaries
tract <- load_data()

#### A. construct panel (some missingness) ------------------------------------

#ensure order of census df matches shapefile
census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]

@@

#establish object for modeling based off of adjusting `panel`
sea_df <- panel %>% 
  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
  arrange(listingQtr, GISJOIN) #arange qtr, tract

#### B. INLA Models of CL Rent ------------------------------------------------

library(INLA)

@@

Home {data-icon="fa-home"}
=====================================

Column
-------------------------------------

### Welcome!

This website provides interactive results for the forecasting models explored in the paper [_Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_](http://github.com/hesscl/smooth-sea/smooth-sea.html).

The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods. The focal series of models regress median one-bedroom rent asked values on different specifications of the panel's correlation structure (i.e. temporal and spatial). All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters. Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e. `r as.yearqtr(Sys.Date())-.25`). The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.

Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic. Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic. You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea).

Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.

_This page was last updated: `r Sys.Date()`_

<br>
<hr>
<br>

**Table of Contents**

Page                |  Description
------------------- | ---------------------------------------------------------
Distribution        | density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
Panel Time-Series   | line graphic to show the observed or modeled temporal structure
Spatial Time-Series | series of maps to show observed change across time
Model Fit           | tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data

Column
-------------------------------------

### Observed vs. Smoothed Rent Estimates

```{r index.maps, echo=F}
obs <- sea_f %>%
  mutate(rent = actualRent) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)

pred <- sea_f %>%
  mutate(rent = bym_Med) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)

comb <- bind_rows(obs, pred) %>%
  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(bym_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
comb <- SharedData$new(comb, ~GISJOIN)

avp_map <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = tooltip)) +
  facet_wrap(~ var) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(avp_map, tooltip = "label") %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  highlight(on = "plotly_click")
```

Distribution
=====================================

```{r, echo=F}
gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
  geom_density() +
  theme_minimal() +
  scale_x_continuous(labels = scales::dollar) +
  scale_color_viridis_d(end = .8) +
  xlab("\nMedian 1B Rent") +
  labs(color = "Quarter")

ggplotly(gg_hist, tooltip = c("color", "x")) %>%
  layout(autosize = T)
```

Panel Time-Series
=====================================

```{r temp.lineplot, echo=F}
obs_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = actualRent,
         qtrDev = value-mean(value, na.rm=T),
         model = "A. Observed")

int_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = int_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "B. Quarter FE") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

ns_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = ns_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "C. Non-Spatial") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

nsar1_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = nsar1_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "D. Non-Spatial AR(1)") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

bym_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = bym_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "E. Spatial AR(1)") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

spt_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value =  spt_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "F. Spatiotemporal AR(1)") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

panel_df <- bind_rows(obs_df, int_df, ns_df, nsar1_df, bym_df, spt_df)

panel_df <- SharedData$new(panel_df, ~GISJOIN)

obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
  geom_line(lwd = .25, color = "black", alpha = .5) +
  geom_point() +
  scale_color_gradient2_tableau() +
  scale_y_continuous(labels = scales::dollar) +
  xlab("") +
  ylab("Median 1B Rent\n") +
  theme_minimal() +
  labs(color = "Quarterly\ndeviation")

ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
  layout(autosize = T) %>%
  animation_opts(frame = 1000, easing = "linear", redraw = FALSE) %>%
  animation_button(hide = T) %>%
  highlight()
```

Spatial Time-Series
=====================================

Column {.tabset}
-------------------------------------

### Observed

```{r spat.series.actual, echo=F}
map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
  facet_grid(~ listingQtr) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_obs) %>%
  layout(autosize = T)
```

### Non-Spatial AR(1)

```{r spat.series.nsar1, echo=F}
map_panel_nsar1 <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = nsar1_Med, frame = listingQtr)) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_nsar1) %>%
  layout(autosize = T) %>%
  animation_opts(1000, transition = 0, redraw = FALSE)
```

### Spatial AR(1)

```{r spat.series.bym, echo=F}
map_panel_bym <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med, frame = listingQtr)) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_bym) %>%
  layout(autosize = T) %>%
  animation_opts(1000, transition = 0, redraw = FALSE)
```

### Spatiotemporal AR(1)

```{r spat.series.spt, echo=F}
map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_spt) %>%
  layout(autosize = T) %>%
  animation_opts(1000, transition = 0, redraw = FALSE)
```

Model Fit
=====================================

Column {.tabset}
-------------------------------------

### Model Legend

Model abbr.                   | Description
----------------------------- | -----------------------------------------------
**int**    | Quarter fixed intercept
           | `r format(form.int)`
**ns**     | Non-spatial tract random effect for each tract, quarter fixed intercept
           | `r format(form.ns)`
**nsar1**  | Non-spatial tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
           | `r format(form.nsar1)`
**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
           |  `r format(form.bym)`
**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter, i.i.d. random effect for tract-quarter (space-time interaction)
           | `r format(form.spt)`

### Accuracy and Information Criteria

```{r, echo=F}
knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("rmse")))
```

<br>
<hr>
<br>

```{r, echo=F}
knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("mae")))
```

<br>
<hr>
<br>

```{r, echo=F}
knitr::kable(data.frame("train_test" = "Training",
                        "int_DIC" = m.int$dic$dic,
                        "ns_DIC" = m.ns$dic$dic,
                        "nsar1_DIC" = m.nsar1$dic$dic,
                        "bym_DIC" = m.bym$dic$dic,
                        "spt_DIC" = m.spt$dic$dic))
```

<br>
<hr>
<br>

```{r, echo=F}
knitr::kable(data.frame("train_test" = "Training",
                        "int_WAIC" = m.int$waic$waic,
                        "ns_WAIC" = m.ns$waic$waic,
                        "nsar1_WAIC" = m.nsar1$waic$waic,
                        "bym_WAIC" = m.bym$waic$waic,
                        "spt_WAIC" = m.spt$waic$waic))
```

### Hyperparameters

```{r, echo=F}
knitr::kable(summary(m.nsar1)$hyperpar, caption = "Non-Spatial AR(1)")
```

<br>
<hr>
<br>

```{r, echo=F}
knitr::kable(summary(m.bym)$hyperpar, caption = "Spatial AR(1)")
```

<br>
<hr>
<br>

```{r, echo=F}
knitr::kable(summary(m.spt)$hyperpar, caption = "Spatiotemporal AR(1)")
```

Column {.tabset}
-------------------------------------

### Non-Spatial AR(1)

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = nsar1_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
  xlab("Actual") +
  ylab("Predicted") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, transition = 0, redraw = T)
```

### Spatial AR(1)

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
  xlab("Actual") +
  ylab("Predicted") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, transition = 0, redraw = FALSE)
```

### Spatiotemporal AR(1)

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = spt_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
  xlab("Actual") +
  ylab("Predicted") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, transition = 0, redraw = FALSE)
```

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Web Data"
author: "Chris Hess"
date: "June 2, 2018"
output: html_document
abstract: "Online classified ads are a key source of information about local rental housing markets. This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA). All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e. 2018 Q2). The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
---

```{r setup, include=FALSE}
#packages
library(tidyverse) #need dev version of ggplot
library(forcats)
library(haven)
library(lubridate)
library(haven)
library(zoo)
library(ggthemes)
library(RColorBrewer)
library(sp)
library(spdplyr)
library(geosphere)
library(rgdal)
library(rgeos)
library(sqldf)
library(spdep)
library(viridis)
library(latex2exp)

@@

#go ahead and save separate versions for graphics later
tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
                  layer = "sea_tract_2000",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F) 
tract2010 <- sea_shp

#source data_load()
source("./helpers/load_data.R")

#use sourced function to load listing summaries
tract <- load_data()
```

<br>
<br>

Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of historic population growth. Growing shares of households are unable to buy a home short of "driving till they qualify", and among households who rent a substantial share pay 30%, if not more of their incomes on their housing costs. The rising cost of housing therefore has a direct role in neighborhood processes of "demographic inversion" like urban gentrification and suburbanizing poverty. Census data (i.e. decennial and American Community Survey) have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication schedules prohibit analyzing any market dynamics beyond long-term trends, particularly among smaller areas of aggregation like neighborhoods. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

This study aims to demonstrate why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography through a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub. I start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, I present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. I conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

<br>

## Background

Existing public data sources like the American Community Survey or decennial census theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts. Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e. ACS estimates are usually at least 2 years old, decennial many more). This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator. For already-leased rental housing units that are sampled, the temporal lag generates measurement error such that data on a given building's current asking rents may be underestimated relative to current tenants' payment levels.

For-profit data sources play a role in filling the information gap about rental housing costs, though they tend to possess their own unique shortcomings. These data sources tend to rely on survey methodologies for data collection, and ask landlords in a given market to outline the average cost for units of a given bedroom size or other specification. In exchange for providing detailed information for their database, the rental data provider may give participating landlords and property management companies access to a slice of information generated via accumulated data. These data tend to be rather opaque in their collection methodologies, with larger or corporate-owner properties most likely to appear in these data relative to independently-owned (i.e. mom and pop) rental units.

<br>

## Data

```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/map_point_density.R") #if config, refresh plot pdf
}
```

![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](./output/map_point_density.png)
<br>

### Web scraping with Helena

<br>

### Data Processing

<br>

### Data Description

```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_lineplot.R") #if config, refresh plot pdf
}
```

![Figure 2. Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e. respective 1-Year Estimates for each year period).](./output/rent_lineplot.png)

```{r rent_map_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_map_across.R") #if config, refresh plot pdf
}
```

![Figure 3. Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](./output/rent_map_across_1B.png)

```{r rent_scatter_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_scatter_across.R") #if config, refresh plot pdf
}
```

![Figure 4. Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](./output/rent_scatter_across_1B.png)
<br>

## Models

```{r construct panel (some missingness), echo=FALSE, message=FALSE, warning=FALSE}
#ensure order of census df matches shapefile
census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]

@@

#establish object for modeling based off of adjusting `panel`
sea_df <- panel %>% 
  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
  arrange(listingQtr, GISJOIN) #arange qtr, tract
```

```{r run models, echo=FALSE, message=FALSE, warning=FALSE}
library(INLA)

@@

`r knitr::kable(fit_stats)`

<br>

## Results

<br>

## Discussion

<br>

## References
